---
title: "Explore vole data"
author: "Mark Ruddy"
date: "2017-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(httr)
```


## Introduction

The aim of this R project is to bring the data on water vole evolution, collected during my PhD [@Ruddy:2011], to life again through the use of `tidyverse` and `shiny`. These weren't available during my PhD and I'm interested in seeing how they can be applied to palaeontological data.


## Datasets

The first stage of this work is reading in the data used during my thesis and tidying / exploring it.

I'm using the supplementary data submitted as part of the these and available online at [@Ruddy:2011].



```{r}


```



Enamel 

```{r}

et <- 

glimpse(et.models.gt4.phdcor)

et.models.gt4.phdcor %>% group_by(OTU) %>%
  summarise(count = n())

```

















## References


Ruddy, M (2011) The Western Palaeartic evolution of the water vole, *Arvicola*. https://pure.royalholloway.ac.uk/portal/en/publications/the-western-palaearctic-evolution-of-the-water-vole-arvicola(2153a197-5a98-472c-8645-68e878517829).html


