---
title: "Explore vole data"
author: "Mark Ruddy"
date: "2017-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Introduction

The aim of this R project is to bring the data on water vole evolution, collected during my PhD [@Ruddy:2011], to life again through the use of `tidyverse` and `shiny`. These weren't available during my PhD and I'm interested in seeing how they can be applied to palaeontological data.


## Datasets

The first stage of this work is reading in the data used during my thesis and tidying / exploring it.

```{r}
load(file = "./data/ages.RData")
load(file = "./data/enamel_microns.RData")
load(file = "./data/enamel_models_phdcor.RData")
load(file = "./data/LocAges_phdcor.RData")
```


```{r}
glimpse(LocAges4.et)
str(LocAges9.shapes)
str(LocAges9.et)
```


```{r}
glimpse(et_micron)
```


```{r}
glimpse(et.models.gt4.phdcor)
```


```{r}
glimpse(LocAges.phdcor)

assemblage <- LocAges.phdcor

assemblage %>% group_by(Lineage) %>%
  summarise(count = n())


```




## References




