---
title: "Explore vole data"
author: "Mark Ruddy"
date: "2017-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Introduction

The aim of this R project is to bring the data on water vole evolution, collected during my PhD [@Ruddy:2011], to life again through the use of `tidyverse` and `shiny`. These weren't available during my PhD and I'm interested in seeing how they can be applied to palaeontological data.


## Datasets

The first stage of this work is reading in the data used during my thesis and tidying / exploring it.

I'm using the supplementary data submitted as part of the these and available online at [@Ruddy:2011].


### Enamel thickness

Read in.

```{r read_enamel, message=FALSE}
# Raw enamel thickness measurements in microns
et.u <- read_csv(file = "./data/SDQ_unmodified_individual.csv", col_names = TRUE)

# Allometrically corrected enamel thickness measurements in microns
et.c <- read_csv(file = "./data/SDQ_corrected_individual.csv", col_names = TRUE)

et.u <- et.u %>%
  rename(ID = X1)

et.c <- et.c %>%
  rename(ID = X1)

# glimpse(et.u)
```

Join tables of raw and allometrically corrected enamel thicknesses by ImageID

```{r}
dim(et.u); dim(et.c)

# Drop common columns and join
et <- et.u %>%
  select(-ID, -OTUID, -OTU, -CSize) %>%
  inner_join(et.c, by = "ImageID")

glimpse(et)
```

There are an extra 14 records in the inner joined table than either on their own. Is this an extra ImageID?

```{r}
# Check for duplicate ImageIDs
et.u %>% select(ImageID) %>%
  group_by(ImageID) %>%
  summarise(count = n()) %>%
  filter(count > 1)


et.c %>% select(ImageID) %>%
  group_by(ImageID) %>%
  summarise(count = n()) %>%
  filter(count > 1)

```
The same 7 ImageIDs are duplicated in both tables. Let's investigate them.

```{r}
et.u %>% group_by(ImageID) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  inner_join(et.u, by = "ImageID") %>%
  select(-count)
  
et.c %>% group_by(ImageID) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  inner_join(et.c, by = "ImageID") %>%
  select(-count)
```
The other variables are duplicated too so we can go ahead and remove one of the duplicates for each ImageID by using ID as a key.

```{r}
# Using ID (record ID) value as basis for selection
# Select single record from each duplicate
dups <- et.u %>% group_by(ImageID) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  inner_join(et.u, by = "ImageID") %>%
  select(ImageID, ID) %>%
  ungroup(ImageID) %>%
  group_by(ImageID) %>%
  top_n(-1, ID)


# Drop the dups ID from the original et tables

et.u.clean <- et.u %>% anti_join(dups, by = "ID")
et.c.clean <- et.c %>% anti_join(dups, by = "ID")

```

Tidy datasets

```{r}
# Need to define separate variables for raw and corrected measurements
glimpse(et.u.clean)

et.u.tidy <- et.u.clean %>% 
  gather(key = wall, value = raw_thick, B1a:L4p) %>%
  select(-SDQ7i)

et.c.tidy <- et.c.clean %>% 
  gather(key = wall, value = allo_thick, B1a:L4p) %>%
  select(-SDQ7i)

```

Join tidy datasets

```{r}
et.u.tidy; et.c.tidy

# Prove to myself that rows are tied together
et.u.tidy %>%
  select(-OTUID, -OTU) %>%
  inner_join(et.c.tidy, by = c("ID" = "ID", "wall" = "wall")) %>%
  mutate(image_diff = ImageID.x - ImageID.y) %>%
  mutate(csize_diff = CSize.x - CSize.y) %>%
  filter(csize_diff != 0)

# Joining tables on 'ID' and 'wall'

et <- et.u.tidy %>%
  select(-OTUID, -OTU, -CSize, -ImageID) %>%
  inner_join(et.c.tidy, by = c("ID" = "ID", "wall" = "wall"))
  
# rm(list = ls(pattern = "et.+|dups"))
```


Re-order columns for convenience
```{r}
et %>%
  select(ID, ImageID, OTUID, OTU, CSize, wall, raw_thick, allo_thick)

```
































## References


Ruddy, M (2011) The Western Palaeartic evolution of the water vole, *Arvicola*. https://pure.royalholloway.ac.uk/portal/en/publications/the-western-palaearctic-evolution-of-the-water-vole-arvicola(2153a197-5a98-472c-8645-68e878517829).html


