---
title: "Explore vole data"
author: "Mark Ruddy"
date: "2017-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Introduction

The aim of this R project is to bring the data on water vole evolution, collected during my PhD [@Ruddy:2011], to life again through the use of `tidyverse` and `shiny`. These weren't available during my PhD and I'm interested in seeing how they can be applied to palaeontological data.

The first stage of this work is reading in the data used during my thesis and tidying / exploring it.

I'm using the supplementary data submitted as part of the these and available online at [@Ruddy:2011].


## Enamel thickness

Read in.

```{r read_enamel, message=FALSE}

# Raw enamel thickness measurements in microns
et.u <- read_csv(file = "./data/SDQ_unmodified_individual.csv", col_names = TRUE)

# Allometrically corrected enamel thickness measurements in microns
et.c <- read_csv(file = "./data/SDQ_corrected_individual.csv", col_names = TRUE)

et.u <- et.u %>%
  rename(ID = X1)

et.c <- et.c %>%
  rename(ID = X1)

# glimpse(et.u)
```

Join tables of raw and allometrically corrected enamel thicknesses by ImageID

```{r}
dim(et.u); dim(et.c)

# Drop common columns and join
et <- et.u %>%
  select(-ID, -OTUID, -OTU, -CSize) %>%
  inner_join(et.c, by = "ImageID")

glimpse(et)
```

There are an extra 14 records in the inner joined table than either on their own. Is this an extra ImageID?

```{r}
# Check for duplicate ImageIDs
et.u %>% select(ImageID) %>%
  group_by(ImageID) %>%
  summarise(count = n()) %>%
  filter(count > 1)


et.c %>% select(ImageID) %>%
  group_by(ImageID) %>%
  summarise(count = n()) %>%
  filter(count > 1)

```
The same 7 ImageIDs are duplicated in both tables. Let's investigate them.

```{r}
et.u %>% group_by(ImageID) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  inner_join(et.u, by = "ImageID") %>%
  select(-count)
  
et.c %>% group_by(ImageID) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  inner_join(et.c, by = "ImageID") %>%
  select(-count)
```
The other variables are duplicated too so we can go ahead and remove one of the duplicates for each ImageID by using ID as a key.

```{r}
# Using ID (record ID) value as basis for selection
# Select single record from each duplicate
dups <- et.u %>% group_by(ImageID) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  inner_join(et.u, by = "ImageID") %>%
  select(ImageID, ID) %>%
  ungroup(ImageID) %>%
  group_by(ImageID) %>%
  top_n(-1, ID)


# Drop the dups ID from the original et tables

et.u.clean <- et.u %>% anti_join(dups, by = "ID")
et.c.clean <- et.c %>% anti_join(dups, by = "ID")

```

Tidy datasets

```{r}
# Need to define separate variables for raw and corrected measurements
glimpse(et.u.clean)

et.u.tidy <- et.u.clean %>% 
  gather(key = wall, value = raw_thick, B1a:L4p) %>%
  select(-SDQ7i)

et.c.tidy <- et.c.clean %>% 
  gather(key = wall, value = allo_thick, B1a:L4p) %>%
  select(-SDQ7i)

```

Join tidy datasets

```{r}
et.u.tidy; et.c.tidy

# Prove to myself that rows are tied together
et.u.tidy %>%
  select(-OTUID, -OTU) %>%
  inner_join(et.c.tidy, by = c("ID" = "ID", "wall" = "wall")) %>%
  mutate(image_diff = ImageID.x - ImageID.y) %>%
  mutate(csize_diff = CSize.x - CSize.y) %>%
  filter(csize_diff != 0)

# Joining tables on 'ID' and 'wall'

et <- et.u.tidy %>%
  select(-OTUID, -OTU, -CSize, -ImageID) %>%
  inner_join(et.c.tidy, by = c("ID" = "ID", "wall" = "wall"))
  
rm(list = ls(pattern = "et.+|dups"))
```


Re-order columns for convenience
```{r}
et <- et %>%
  select(ID, ImageID, OTUID, OTU, CSize, wall, raw_thick, allo_thick)
```

Round 'allo_thick' values to 2 sf.

```{r}
et$allo_thick <- round(et$allo_thick)
```

### Ammendments

OTU names for correction
```{r}
et <- et %>% 
  mutate(OTU = if_else(OTU == 'Pesko "brick red" strata', "Pesko brick red strata", OTU)) %>%
  mutate(OTU = if_else(OTU == 'Trou du Frontal', "Trou du Frontal 1er niveau ossifere", OTU)) %>%
  mutate(OTU = if_else(OTU == 'Miesenheim I', "Miesenheim I C-H", OTU)) %>%
  mutate(OTU = if_else(OTU == 'Kuznetsovka', "Kuznetsovka Bed 3", OTU)) %>%
  mutate(OTU = if_else(OTU == 'Donskaya Negatchevka', "Donskaya Negatchevka 9", OTU)) %>%
  mutate(OTU = if_else(OTU == 'Cudmore Grove', "Cudmore Grove 'Detrital mud'", OTU))

  
  
  
```

Tidy data export

```{r}
write_csv(x = et, path = "./data/enamel_thickness.csv", col_names = TRUE)
```

## Sites


Read in.

```{r read_sites, message=FALSE}

# Fossil and modern sites
fossil <- read_csv(file = "./data/FossilLocalities.csv", col_names = TRUE)
modern <- read_csv(file = "./data/RecentLocalities.csv", col_names = TRUE)
fossil.asmb <- read_csv(file = "./data/FossilAssemblages.csv", col_names = TRUE)
```

### Ammendments

Some corrections to entries in the dfs. 

Two missing archive values that should be NHML
```{r}
modern <- modern %>% mutate(Archives = if_else(Locality == "Welsh Borders", "NHML", Archives)) %>%
  mutate(Archives = if_else(Locality == "Central Scotland", "NHML", Archives))
```

Fossil assemblage name
```{r}
fossil.asmb <- fossil.asmb %>% 
  mutate(Assemblage = if_else(Assemblage == "Vindija E", "Vindija E-F", Assemblage)) %>% 
  mutate(Assemblage = if_else(Assemblage == "Tonchesberg II B", "Tonchesburg II 11-13", Assemblage)) %>%
  mutate(Assemblage = if_else(Assemblage == "Cudmore Grove `Detrital mud'", "Cudmore Grove 'Detrital mud'", Assemblage))
  
  
```


### Names and coordinates

Modern sites have no assemblage name or age and fossil assemblages have no coordinates. These are needed for querying. Need to combine elements of different tables to rectify this.

Make sites df by combining fossil assemblages and modern sites. Add age columns to modern sites, then adding fossil assemblage long/lats and site names to fuzzy matching site name to assemblage name.

```{r}

# Add Min_age and Max_age to modern
modern$Min_age <- 0
modern$Max_age <- 0

# Change age col names in fossil assemblage
fossil.asmb <- fossil.asmb %>% rename(Min_age = `Min age`, Max_age = `Max age`)

# Add lat/long columns to fossil assemblage
fossil.asmb$Latitude <- NA
fossil.asmb$Longitude <- NA

# Drop n columns
modern <- modern %>% select(-n)
fossil.asmb <- fossil.asmb %>% select(-n)

# Combine
sites <- bind_rows(modern, fossil.asmb)

# Map modern Locality to Assemblage
sites <- sites %>% mutate(Assemblage = if_else(is.na(Assemblage), Locality, Assemblage))

# Change ages from years to thousands of years 

sites <- sites %>% 
  mutate(Min_age = round(.$Min_age * 10^-3, digits = 2)) %>%
  mutate(Max_age = round(.$Max_age * 10^-3, digits = 2))
  
# rm(modern, fossil)

```



Now do fuzzy join between assemblage name and fossil site name to add site name to assemblage df
```{r}
# install.packages("fuzzyjoin")
library(fuzzyjoin)

sites.test <- sites %>% 
  stringdist_full_join(fossil, by = c("Assemblage" = "Locality"), method = "lcs", max_dist = 4, distance_col = "dist")

rm(sites.test)
```

Best performing fuzzy match is not great. Let's try match based on truncated site name/assemblage name then update based on that match.


```{r}
library(stringr)

# Create joining columns in both dfs. Create common key.
sites.match <- sites %>% mutate(siteID = str_trunc(Assemblage, width = 5, side = "right", ellipsis = ""))
fossil.match <- fossil %>% mutate(siteID = str_trunc(Locality, width = 5, side = "right", ellipsis = ""))
# Sub-select columns of interest from fossil df
fossil.match <- fossil.match %>% select(Locality, siteID, Longitude, Latitude)

# Join sites and fossil on new siteID
sites.match <- sites.match %>% full_join(fossil.match, by = c("siteID" = "siteID"))

# Update Locality.x, Long....x and Lat....x from .y columns
sites.match <- sites.match %>% mutate(Locality.x = ifelse(is.na(Locality.x), Locality.y, Locality.x)) %>%
  mutate(Longitude.x = if_else(is.na(Longitude.x), Longitude.y, Longitude.x)) %>%
  mutate(Latitude.x = if_else(is.na(Latitude.x), Latitude.y, Latitude.x)) %>%
  select(-siteID, -Locality.y, -Longitude.y, -Latitude.y) %>%
  rename(Locality = Locality.x, Latitude = Latitude.x, Longitude = Longitude.x)

sites <- sites.match

```

Drop some Assemblages not present in et table and drop X1 column
```{r}
sites <- sites %>% 
  filter(Assemblage != "Regourdou 3") %>%
  filter(Assemblage != "Komarowa D") %>%
  filter(Assemblage != "Kalman Lambrecht IV") %>%
  filter(Assemblage != "Ightham Fissures") %>%
  filter(Assemblage != "Horvolgy") %>%
  filter(Assemblage != "Divje babe I Facies A") %>%
  filter(Assemblage != "Bisnik 12") %>%
  filter(Assemblage != "Brillenhohle IV") %>%
  filter(Assemblage != "Bridged Pot inner slope") %>%
  filter(Assemblage != "Biedensteg II") %>%
  filter(Assemblage != "Balla") %>%
  filter(Assemblage != "St. Bertrand de Comminge") %>%
  filter(Assemblage != "Ormozu") %>%
  filter(Assemblage != "Huesca") %>%
  filter(Assemblage != "Court St Etienne") %>%
  filter(Assemblage != "Casian") %>%
  select(-X1)
```


```{r}
# Clean up

rm(sites.match, fossil, fossil.asmb, fossil.match, modern)

```

Test join between et and sites dfs. Should be no records missing from join.
```{r}
sites.test <- sites %>% anti_join(et, by = c("Assemblage" = "OTU"))
rm(sites.test)
```

Tidy data export

```{r}
write_csv(x = sites, path = "./data/enamel_sites.csv", col_names = TRUE)
```





















## References


Ruddy, M (2011) The Western Palaeartic evolution of the water vole, *Arvicola*. https://pure.royalholloway.ac.uk/portal/en/publications/the-western-palaearctic-evolution-of-the-water-vole-arvicola(2153a197-5a98-472c-8645-68e878517829).html


